Batch Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-02-09 paper/background.tex
Paper update

We handle the scenarios where the topology of the analog circuit is fixed. This
is practical as there are usually a lot of classical topologies for a given
design task. Once the circuit topology is fixed, the designer has to choose the
appropriate design parameters according to the specifications and the circuit
device model. What we want to do is automatically searching for the optimal
design parameters. This problem can then be formulated as a bound-constrained
black-box optimization problem:
\begin{equation}
    \label{eq:Formulation}
    \text{minimize}~\mathrm{FOM}(\bm{x}),
\end{equation}
where $\bm{x} \in \textrm{D}$ is the vector of design variables,
$\mathrm{FOM}(\bm{x})$ is the objective constructed from the design
specifications, the $\mathrm{FOM}(\bm{x})$ can be deterministric or noisy
depending on the design specifications. Given the design parameters $\bm{x}$,
the FOM value can be obtained by commercial circuit simulators like HSPICE or
Spectre.

@@

The objective function $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} can be
approximated by Gaussian process (GP) model~\cite{GPML}. The GP model is the
most commonly used model for Bayesian optimization. The advantage of GP is that
it provides a well-calibrated uncertainty of prediction. GP is characterized by
a mean function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In
this work, we use squared-exponential ARD kernel~\cite{GPML}, and a constant
mean function $m(\bm{x}) = \mu_0$ for all our experiments. By default, we
assume the objective function evaluations are influenced by i.i.d. noise
$\epsilon_t \sim N(0, \sigma_n^2)$ and set the noise level $\sigma_n^2$ as a
hyperparameter.  The introduction of the i.i.d noise also helps to improve the
numerical stability.

Given a new data point $\bm{x}$, the prediction of $f(\bm{x})$ is
not a scalar value, but a predictive distribution
\begin{equation}
f(\bm{x}) \sim N(\mu(\bm{x}),
\sigma^2(\bm{x})),
\label{eq:GPRPred}
\end{equation}
where $\mu(\bm{x})$ and $\sigma^2(\bm{x})$ can be expressed as
\begin{equation}
        \begin{array}{lll}
            \mu(\bm{x})      &=& \mu_0 + k(\bm{x},X)[K+\sigma_n^2I]^{-1}(\bm{y} - \mu_0) \\
            \sigma^2(\bm{x}) &=& k(\bm{x}, \bm{x}) - k(\bm{x}, X)[K+\sigma_n^2]^{-1}k(X, \bm{x}),
        \end{array}
    \label{eq:GPRPredEqNoisy}
\end{equation}
where $k(\bm{x}, X) = (k(\bm{x}, \bm{x}_1), \dots, k(\bm{x},
\bm{x}_N))^T$ and $k(X, \bm{x}) = k(\bm{x}, X)^T$. The
$\mu(\bm{x})$ can be viewed as the prediction of the function value, while
the $\sigma^2(\bm{x})$ is a measure of uncertainty of the prediction.

@@

The LCB function is defined as follows:
\begin{equation}
    \label{eq:LCB}
    \mathrm{LCB}(\bm{x}) = \mu(\bm{x}) - \kappa \sigma(\bm{x}),
\end{equation}
where the $\mu(\bm{x})$ and the $\sigma(\bm{x})$ are the predictive value and uncertainty of GP, $\kappa$ is a parameter that balances the exploitation and exploration.

>>>>

When designing integrated circuits, the designers have to decide what circuit
topology to use and then set the corrsponding design parameters. In this work,
we handle the scenarios where the topology of the analog circuit is fixed. This
is practical as there are usually a lot of classical topologies for a given
design task, so unlike digital circuits, choosing appropriate topology is
relatively easy.

Once the circuit topology is fixed, the designer has to choose the appropriate
design parameters according to the specifications and the circuit device model.
What we want to do is automatically searching for the optimal design
parameters. This problem can then be formulated as a bound-constrained
black-box optimization problem:
\begin{equation}
    \label{eq:Formulation}
    \text{minimize}~\mathrm{FOM}(\bm{x}),
\end{equation}
where $\bm{x} \in \textrm{D}$ is the vector of design variables,
$\mathrm{FOM}(\bm{x})$ is the objective constructed from the design
specifications, the $\mathrm{FOM}(\bm{x})$ can be deterministric or noisy
depending on the design specifications. Given the design parameters $\bm{x}$,
the FOM value can be obtained by commercial circuit simulators like HSPICE or
Spectre.

@@

The objective function $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} can be
approximated by Gaussian process (GP) model~\cite{GPML}. The GP model is the
most commonly used model for Bayesian optimization. The advantage of GP is that
it provides a well-calibrated uncertainty of prediction. GP is characterized by
a mean function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In
this work, we use squared-exponential ARD kernel~\cite{GPML}, and a constant
mean function $m(\bm{x}) = \mu_0$ for all our experiments. By default, we
assume the objective function evaluations are influenced by i.i.d. noise
$\epsilon_t \sim N(0, \sigma_n^2)$ and set the noise level $\sigma_n^2$ as a
hyperparameter.  The introduction of the i.i.d noise also helps to improve the

Denote the training set as $\{X, \bm{y}\}$ where $X = \{\bm{x}_1, \dots, \bm{x}_N\}$ and $\bm{y} = \{y_1, \dots, y_N\}$, given a new data point $\bm{x}$, the prediction of $f(\bm{x})$ is not a scalar value, but a predictive distribution
\begin{equation}
f(\bm{x}) \sim N(\mu(\bm{x}),
\sigma^2(\bm{x})),
\label{eq:GPRPred}
\end{equation}
where $\mu(\bm{x})$ and $\sigma^2(\bm{x})$ can be expressed as
\begin{equation}
        \begin{array}{lll}
            \mu(\bm{x})      &=& \mu_0 + k(\bm{x},X)[K+\sigma_n^2I]^{-1}(\bm{y} - \mu_0) \\
            \sigma^2(\bm{x}) &=& k(\bm{x}, \bm{x}) - k(\bm{x}, X)[K+\sigma_n^2I]^{-1}k(X, \bm{x}),
        \end{array}
    \label{eq:GPRPredEqNoisy}
\end{equation}
where $k(\bm{x}, X) = (k(\bm{x}, \bm{x}_1), \dots, k(\bm{x},
\bm{x}_N))^T$ and $k(X, \bm{x}) = k(\bm{x}, X)^T$. The
$\mu(\bm{x})$ can be viewed as the prediction of the function value, while
the $\sigma^2(\bm{x})$ is a measure of uncertainty of the prediction.

@@

The LCB function is defined as follows:
\begin{equation}
    \label{eq:LCB}
    \mathrm{LCB}(\bm{x}) = \mu(\bm{x}) - \kappa \sigma(\bm{x}),
\end{equation}
where the $\mu(\bm{x})$ and the $\sigma(\bm{x})$ are the predictive value and uncertainty of GP defined in \eqref{eq:GPRPredEqNoisy}, $\kappa$ is a parameter that balances the exploitation and exploration.

### 2018-02-09 paper/mace.tex
Paper update

Each acquisition function represents a unique selection strategy, different acquisition functions may not agree with each other about where to sample the next point. For example, the value of LCB function always decreases as the $\sigma(\bm{x})$ increases. However, for the PI function, when $\sigma(\bm{x})$ increases, the value of PI would decrease when $\mu(\bm{x}) < \tau$, and increase when $\mu(\bm{x}) > \tau$. For the EI function, the values of EI function at already sampled points would always be lower than any unsampled locations, while this property does not hold for the LCB function.

@@

The proposed MACE algorithm is described in Algorithm~\ref{alg:MACE}.

\begin{algorithm}
    \caption{Multi-objective Acquisition Ensemble Algorithm}
    \label{alg:MACE}
    \begin{algorithmic}[1]
        \REQUIRE Number of initial sampling points $N_{init}$, number of iterations $N_{iter}$, batch size $B$.
        \STATE Randomly sample $N_{init}$ points in the design space
        \STATE Construct initial GP model
        \FOR{t = 1, 2, \dots, $N_{iter}$}
        \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
        \STATE Find the Pareto front of LCB, EI, PI functions using the DEMO algorithm
        \STATE Randomly sample $B$ points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto-optimal points
        \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
        \STATE Update the GP model
        \ENDFOR
        \STATE \textbf{Return} best $f(\bm{x})$ recorded during iterations
    \end{algorithmic}
\end{algorithm}

@@

We illustrate the proposed MACE algorithm using an example of a real-world amplifier circuit. The optimization objective is to maximize the phase margin (PF) of the amplifier, so the FOM is defined as $\mathrm{FOM}(\bm{x}) = - \mathrm{PF}(\bm{x})$. The width of one of its transistor is the design variable. We sweep the width of the transistor and perform HSPICE simulations to get the FOM values. The curve of FOM values is plotted in Figure~\ref{fig:PF_example_1} (the blue line). Several points are randomly sampled from the FOM curve to train the GP model. The LCB, EI, PI functions and the Pareto front of the acquisition functions are plotted in Figure~\ref{fig:PF_example_2}. We can see from Figure~\ref{fig:PF_example_2} that the optimal locations of the three acquisition functions are different, while their best trade-off is captured by the Pareto front. The Pareto set that represents the best trade-off between the three acquisition functions is the interval $[43, 50.4]$, as plotted in Figure~\ref{fig:PF_example_1}. The candidate points for the next batch of evaluations are randomly sampled from the Pareto set.

% \textcolor{red}{TODO: If there is space, discuss the approximation of EI in a new subsection "Implementation details"}

>>>>

Each acquisition function represents a unique selection strategy, different acquisition functions may not agree with each other about where to sample the next point. For example, the value of LCB function always decreases as the $\sigma(\bm{x})$ increases. However, for the PI function, when $\sigma(\bm{x})$ increases, the value of PI would decrease when $\mu(\bm{x}) < \tau$, and increase when $\mu(\bm{x}) > \tau$. For the EI function, the values of EI function at already sampled points would always be lower than the values at any unsampled locations, while this property does not hold for the LCB function.

@@

@@

% \textcolor{red}{TODO: If there is space, discuss the approximation of EI in a new subsection "Implementation details"}
The proposed MACE algorithm is described in Algorithm~\ref{alg:MACE}.

\begin{algorithm}[!htb]
    \caption{Multi-objective Acquisition Ensemble Algorithm}
    \label{alg:MACE}
    \begin{algorithmic}[1]
        \REQUIRE Number of initial sampling points $N_{init}$, number of iterations $N_{iter}$, batch size $B$.
        \STATE Randomly sample $N_{init}$ points in the design space
        \STATE Construct initial GP model
        \FOR{t = 1, 2, \dots, $N_{iter}$}
        \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
        \STATE Find the Pareto front of LCB, EI, PI functions using the DEMO algorithm
        \STATE Randomly sample $B$ points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto-optimal points
        \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
        \STATE Update the GP model
        \ENDFOR
        \STATE \textbf{Return} best $f(\bm{x})$ recorded during iterations
    \end{algorithmic}
\end{algorithm}

In Figure~\ref{fig:PF_example}, we illustrate the proposed MACE algorithm using an example of a real-world amplifier circuit. The optimization objective is to maximize the phase margin (PF) of the amplifier, so the FOM is defined as $\mathrm{FOM}(\bm{x}) = - \mathrm{PF}(\bm{x})$. The width of one of its transistor is the design variable. We sweep the width of the transistor and perform HSPICE simulations to get the FOM values. The curve of FOM values is plotted in Figure~\ref{fig:PF_example_1} (the blue line). Several points are randomly sampled from the FOM curve to train the GP model. The LCB, EI, PI functions and the Pareto front of the acquisition functions are plotted in Figure~\ref{fig:PF_example_2}. We can see from Figure~\ref{fig:PF_example_2} that the optimal locations of the three acquisition functions are different, while their best trade-off is captured by the Pareto front. The Pareto set that represents the best trade-off between the three acquisition functions is the interval $[43, 50.4]$, as plotted in Figure~\ref{fig:PF_example_1}. The candidate points for the next batch of evaluations are randomly sampled from the Pareto set.
